---
title: "03_DescriptiveReport_Analysis_26-11-2022"
author: "Mehak Khanna"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA, warning = FALSE, message=FALSE)
library(tidyverse)
library(tableone)
library(table1)
library(here)
library(lubridate)
library(readxl)
library(eeptools)
library(knitr)
scc_doi<-read_csv(here("Data","clean_scc_doi_26-11-2022.csv"))

baseline_scc_doi<- read_csv(here("Data","baseline_scc_doi_26-11-2022.csv"))

```


```{r DOI category tables, include=T, echo=FALSE}
##True DOI and tumour thickness----------------
Tab1<-CreateCatTable(vars = "doi_cat",
               data = baseline_scc_doi,
               includeNA = T)%>% print(showAllLevels=T)
kableone(Tab1, caption = "1: Frequency of categories of DOI")

Tab2<-CreateCatTable(vars = c("know_doi_124", "surgery_type"),
                     strata = "doi_cat",
                     data = baseline_scc_doi, includeNA = T, test = F)%>% print(showAllLevels=T)
kableone(Tab2, caption= "2: participants whose DOI is known and type of surgey VS the category of DOI recorded for them")

#Demographic-------------------
Tab3<-CreateTableOne(vars = c("age", "GenderSexe", "tobacco_hist", "tobacco_stat_diag", "alc_hist", "alc_stat_diag"),data = baseline_scc_doi, includeNA = T)%>%
  print(showAllLevels=TRUE)

kableone(Tab3, caption="3: Descriptive table of Demgraphic variables")

##Length of follow-up---------------
class(scc_doi$Follow_up_Date)

kable(scc_doi%>%
  group_by(Patient.IDID.du.patient)%>%
  summarise(min(Follow_up_Date, na.rm=T), max(Follow_up_Date, na.rm = T),
            len_flw_mon= as.numeric(`max(Follow_up_Date, na.rm = T)`-`min(Follow_up_Date, na.rm = T)`)/(365.25/12))%>%
  summarise(mean=mean(len_flw_mon), SD=sd(len_flw_mon), 
            median= median(len_flw_mon), min=min(len_flw_mon),
            max=max(len_flw_mon),IQR = IQR(len_flw_mon)), 
  caption = "4: Summary of lenth of follow-up in Months", digits = 2, format = "simple") 

```


Length of follow-up (in months) is at present calculated till the last date provided in the dataset. This means for some patients even after recurrence or metastasis the follow-ups continued. 
The outcome of our study is recurrence, once the outcome has occurred should we continue continuing the follow up time?


Below is the rounding to nearest whole number.
Please note in variable 42 there is 0.74 that gets assigned to 1 on the other hand,

```{r, echo=FALSE}
#Depth of Invasion----------------
table(baseline_scc_doi$doi_42, round(baseline_scc_doi$doi_42_round), exclude = NULL)
```

variable 125 has 0.36 that gets assigned 0. 
```{r, echo=FALSE}
table(baseline_scc_doi$doi_125, round(baseline_scc_doi$doi_125), exclude = NULL)
```
I'm using varibale 125 against DOI category to provide you tables. (Please let me know if that is alright)


Below is a summary table for TRUE DOI and Tumor thickness
```{r, echo=FALSE}
Tab5<-CreateContTable(vars = "doi_125",
                strata = "doi_cat",
                data = baseline_scc_doi,
                funcNames = c("mean","sd","Median","min","max"),
                test = F, smd = F)

print(summary(Tab5))

```

```{r, echo=FALSE}
baseline_scc_doi$doi_125_round<-as.factor(round(baseline_scc_doi$doi_125))

Tab6<-CreateCatTable(vars = "doi_125_round",
               strata = "doi_cat",
               data = baseline_scc_doi,
               test = F, includeNA = T)%>% print(showAllLevels=T)
kableone(Tab6, caption= "6: recorded DOI in mm by DOI category")

Tab7<-CreateCatTable(vars = "doi_125_round",
               strata = c("doi_cat","surgery_type"),
               data = baseline_scc_doi,
               test = F, includeNA = T)%>% print(showAllLevels=T)
kableone(Tab7, caption= "7: recorded DOI in mm by DOI category and further separated by type of Surgery")


tab8<-CreateTableOne(vars = "primary_site",
               data = baseline_scc_doi, includeNA = T )%>%
  print(ShowAllLevels=TRUE)

kableone(tab8, caption="8: Distribution of site of primary lesion")


tab9<-CreateTableOne(vars = c("t_clinc_stage", "n_clinc_stage","m_clinc_stage"),
               data = baseline_scc_doi,
               includeNA = T)%>% print(ShowAllLevels=T)

kableone(tab9, caption="9: Distribution clinical TNM staging of primary lesion")


tab10<- CreateCatTable(vars = "n_clinc_stage",
                       strata = "n_radio_stage",
                       data = baseline_scc_doi, includeNA = T, test = F)%>% print(showAllLevels=T)
kableone(tab10, caption= "10: Probability fo clinic exam differing from radiographic results")


tab11<-CreateTableOne(vars = "Pathological.stage.overall....Stage.Pathologique.Global",
                     data = baseline_scc_doi,
                     includeNA = T)%>%print(ShowAllLevels=T)

kableone(tab11, caption="11: Distribution overall Pathologic staging of tumor")
```

```{r, Echo=FALSE}
tab12<-CreateTableOne(vars = "tab_n_clinc_stage",
               strata = "tab_n_patho_stage",
               data = baseline_scc_doi,
               includeNA = T, test = F) %>% print(showAllLevels=T)
kableone(tab12, caption="12: Clinical node stages stratified by pathological node stage " )

baseline_scc_doi<-baseline_scc_doi%>%
  mutate(tab_n_clinc_radio_stage = case_when((tab_n_clinc_stage == "N0" & tab_n_radio_stage == "N0")~ "N0_clinc_radio",
                                             (tab_n_clinc_stage != "N0" | tab_n_radio_stage != "N0") ~ "All not N0_clinc_radio"))

Tab13<-CreateCatTable(vars = "tab_n_clinc_radio_stage",
               strata = "tab_n_patho_stage",
               data = baseline_scc_doi,includeNA = T, test = F)%>% print(showAllLevels=T)
kable(Tab13, caption= "13: Clinical and Radiographic Nodes combination againist Pathologic Node")


tab14<-CreateTableOne(vars = c("biopsy_differen",
                              "final_patho_differen"),
               data = baseline_scc_doi,
               includeNA = T)%>% print(showAllLevels=T)
kableone(tab14, caption = "14: distribution of participants by degree of differentiation")

baseline_scc_doi$Num_positive_nodes_fac<-as.factor(baseline_scc_doi$Num_positive_nodes_fac)
tab15<-CreateTableOne(vars = "Num_positive_nodes_fac",
               data= baseline_scc_doi,
               includeNA = T) %>% print(showAllLevels=T)
kableone(tab15, caption="15: Distribution by no. of positive nodes")


tab16<-CreateTableOne(vars = c("margins","neg_margins","extra_capsu","peri_neu_inva",
                        "vas_lymph_inv","bone_inv","p16_status"),
               data = baseline_scc_doi,
               includeNA = T) %>% print(showAllLevels=T)
kableone(tab16, caption="16: Distribution of participants by margin of tumor, perinural invasion and etc")
```

```{r, echo=FALSE}

tab17a<-CreateContTable(vars = "age",
                       funcNames = c("n","mean","sd","median","min","max"),
                       strata = "reason_death",
                      data = (scc_doi%>%filter(...148 == "Expired")),
                      test = F)
print(summary(tab17a))


tab17<-CreateCatTable(vars = c("GenderSexe", "tobacco_hist", "tobacco_stat_diag", "alc_hist", "alc_stat_diag","t_clinc_stage","n_clinc_stage", "t_patho_stage","n_patho_stage", "Pathological.stage.overall....Stage.Pathologique.Global","surgery_type","Num_positive_nodes_fac","Pathological.stage.overall....Stage.Pathologique.Global","Was.there.Radiotherapy...Une.Radiotherapie.a.t.elle.administree..","Was.there.Chemotherapy...Une.Chimiotherapie.a.t.elle.administree.."),
                      strata = "reason_death",
                      data = (scc_doi%>%filter(...148 == "Expired")),
                      includeNA = T, test = F)%>% print(showAllLevels=T)
kableone(tab17, caption= "17: Distribution of participant characteristics by reason of death for those who expired")

```

