---
title: "Predictors of occult lymph node metastasis"
author: "Mehak Khanna"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA, warning = FALSE, message=FALSE)
library(tidyverse)
library(tableone)
library(table1)
library(here)
library(readr)
library(lubridate)
library(readxl)
library(eeptools)
library(knitr)

baseline_scc_doi<- read_csv(here("Data","clean_baseline_05-12-2022.csv"))
```

## Univariate logistic regression  - Model 1
<br/>
1) Predicting probability of occult metastasis in lymph nodes
<br/>
2) risk factor - Depth of Invasion (rounded)

```{r DOI model, echo=FALSE}
class(baseline_scc_doi$doi_125_round)
baseline_scc_doi$doi_125_round<-factor(baseline_scc_doi$doi_125_round)

# baseline_scc_doi%>%
#   #select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, n_patho_stage, n_clinc_stage)%>%
#   filter(doi_cat== "true_doi" & !is.na(doi_125_round) & n_clinc_stage== "N0")%>% 
#   count(n_patho_stage) #leaves us with 82 participants
# 
# baseline_scc_doi%>%
#   select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, n_patho_stage, n_clinc_stage)%>%
#   filter(doi_cat== "true_doi" & !is.na(doi_125_round) & n_clinc_stage== "N0")%>% 
#   group_by(doi_125_round)%>% arrange(doi_125_round) %>%view() #add filter of removing those for whom path_n stage not applicable
#   
# baseline_scc_doi%>%
#   select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, n_patho_stage, n_clinc_stage)%>%
#   filter(doi_cat== "true_doi" & !is.na(doi_125_round) & n_clinc_stage== "N0" & n_patho_stage != "Not Applicable")%>% 
#   count(n_patho_stage)

log.reg<-baseline_scc_doi%>%
  select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, doi_125,n_patho_stage, n_clinc_stage)%>%
  filter(doi_cat== "true_doi" & !is.na(doi_125_round) & n_clinc_stage== "N0" & n_patho_stage != "Not Applicable")%>% 
  group_by(doi_125_round)%>% arrange(doi_125_round) #%>% view()

log.reg$occultmet<-NA
for (i in 1:nrow(log.reg)) {
  if (log.reg$n_patho_stage[i] != log.reg$n_clinc_stage[i]){
    log.reg$occultmet[i]<- 1 #meaning yes occult metastasis seen
  } else{
    log.reg$occultmet[i]<-0 #meaning no occult metastasis
  }
}

#ploting the two variables before logistic regression
#class(log.reg$doi_125_round)
log.reg$doi_125_round_num<-as.numeric(as.character(log.reg$doi_125_round))#changing back to numeric
table(log.reg$occultmet, log.reg$doi_125_round_num, exclude = NULL)
#table(log.reg$doi_125_round, log.reg$doi_125_round_num, exclude = NULL)
boxplot(doi_125_round_num ~ occultmet, data = log.reg, xlab = "Occult metastasis", ylab = "DOI",
        main = "Distribution of occult metastasis by DOI for patients who reported TRUE_DOI and clinic stage was N0") # few outliers


##Models predicting occult metastasis--------------------
##DOI
#fitting the model
lr.fit<-glm(occultmet ~ doi_125_round_num, data = log.reg, family = binomial)
summary(lr.fit)

#getting odds ratio with confidence interval
round(exp(lr.fit$coefficients), digits = 2)
round(exp(confint(lr.fit)), digits = 3)
#odds ratio for DOI is 1.02. This means the odds of occult metastasis increase by 2% for each mm increase in DOI
```
###Interpretation of the results:
Occult metastasis ~ Depth of invasion(rounded)
<br/>
OR = 1.02 (95% CI 0.95 - 1.09)
As the depth of invasion increases by 1mm the odds of occult metastasis increases by 2%
<br/>
This result is not statistically significant as the null value of 1 is included in the 95% CI.


```{r DOI model best fit line, echo=FALSE}
#getting predicted values
predicted_data<-data.frame(doi_125_round_num = seq(min(log.reg$doi_125_round_num), 
                                                   max(log.reg$doi_125_round_num),
                                                   len = 500))

predicted_data$occultmet<-predict(lr.fit, predicted_data, type = "response") #geting the predicted values
  
predit_se<-predict(lr.fit,  predicted_data, type = "response", se.fit = T) #se.fitting to get CI for predicted values

predicted_data$lowerCI<-predit_se$fit-1.96*predit_se$se.fit
predicted_data$upperCI<-predit_se$fit+1.96*predit_se$se.fit


ggplot(data = log.reg,
       aes(x=doi_125_round_num,
           y=occultmet))+
  geom_point()+
  stat_smooth(method = "glm", se =TRUE, method.args = list(family=binomial))+
  xlab(label = "Depth of invasion(mm)")+
  ylab(label = "Probability fo occult lymphnode metstasis")+
  labs(title= "Best fit line for predicting occult metastasis based on depth of invasion of tumor")+
  theme_light()
```

## Univariate logistic regression  - Model 2
<br/>
1) Predicting probability of occult metastasis in lymph nodes
<br/>
2) risk factor - Depth of Invasion (decimal)


```{r, echo=FALSE}
log.reg_doi<-baseline_scc_doi%>%
  select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125, n_patho_stage, n_clinc_stage)%>%
  filter(doi_cat== "true_doi" & !is.na(doi_125) & n_clinc_stage== "N0" & n_patho_stage != "Not Applicable")%>% 
  group_by(doi_125)%>% arrange(doi_125) %>% view()

log.reg_doi$occultmet<-NA
for (i in 1:nrow(log.reg_doi)) {
  if (log.reg_doi$n_patho_stage[i] != log.reg_doi$n_clinc_stage[i]){
    log.reg_doi$occultmet[i]<- 1 #meaning yes occult metastasis seen
  } else{
    log.reg_doi$occultmet[i]<-0 #meaning no occult metastasis
  }
}

#class(log.reg_doi$doi_125)
#boxplot(doi_125 ~ occultmet, data = log.reg_doi, xlab = "Occult metastasis", ylab = "DOI_cont",
        #main = "Distribution of occult metastasis by DOI for patients who reported TRUE_DOI and clinic stage was N0") # few outliers

#fitting the model
lr.fit_doicont<-glm(occultmet ~ doi_125, data = log.reg_doi, family = binomial)
summary(lr.fit_doicont)
confint(lr.fit_doicont)


#getting odds ratio with confidence interval
round(exp(lr.fit_doicont$coefficients), digits = 2)
round(exp(confint(lr.fit_doicont)), digits = 2)
#odds ratio for DOI is 1.02. This means the odds of occult metastasis increase by 2% for each mm increase in DOI

```

###Interpretation of the results:
Occult metastasis ~ Depth of invasion(decimal)
<br/>
OR = 1.02 (95% CI 0.96 - 1.1)
As the depth of invasion increases by 1mm the odds of occult metastasis increases by 2%
<br/>
This result is not statistically significant as the null value of 1 is included in the 95% CI.

```{r, echo=FALSE}
#getting predicted values
predicted_data_doicont<-data.frame(doi_125 = seq(min(log.reg_doi$doi_125), 
                                                   max(log.reg_doi$doi_125),
                                                   len = 500))

predicted_data_doicont$occultmet<-predict(lr.fit_doicont, predicted_data_doicont, type = "response") #getting the predicted values

predit_doicon_se<-predict(lr.fit_doicont,  predicted_data_doicont, type = "response", se.fit = T) #se.fitting to get CI for predicted values

predicted_data_doicont$lowerCI<-predit_doicon_se$fit-1.96*predit_doicon_se$se.fit
predicted_data_doicont$upperCI<-predit_doicon_se$fit+1.96*predit_doicon_se$se.fit


ggplot(data = log.reg_doi,
       aes(x=doi_125,
           y=occultmet))+
  geom_point()+
  stat_smooth(method = "glm", se =TRUE, method.args = list(family=binomial))+
  xlab(label = "Depth of invasion(mm)-continous: decimal point")+
  ylab(label = "Probability fo occult lymphnode metstasis")+
  theme_light()
```


## Univariate logistic regression  - Model 3
<br/>
1) Predicting probability of occult metastasis in lymph nodes
<br/>
2) risk factor - Perinural Invasion

```{r , echo=FALSE}
# baseline_scc_doi%>%
#   #select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, n_patho_stage, n_clinc_stage)%>%
#   filter(n_clinc_stage== "N0")%>% 
#   count(n_patho_stage) #leaves us with 82 participants


PNI_log.reg<-baseline_scc_doi%>%
  select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, peri_neu_inva,n_patho_stage, n_clinc_stage)%>%
  filter(n_clinc_stage== "N0" & n_patho_stage != "Not Applicable", !is.na(peri_neu_inva))%>% view()
#127 participants for this model

PNI_log.reg$occultmet<-NA
for (j in 1:nrow(PNI_log.reg)) {
  if (PNI_log.reg$n_patho_stage[j] != PNI_log.reg$n_clinc_stage[j]){
    PNI_log.reg$occultmet[j]<- 1 #meaning yes occult metastasis seen
  } else{
    PNI_log.reg$occultmet[j]<-0 #meaning no occult metastasis
  }
}  

table(PNI_log.reg$occultmet, PNI_log.reg$peri_neu_inva, exclude = NULL)

pni.lr.fit<-glm(occultmet~peri_neu_inva, data = PNI_log.reg, family = binomial)
summary(pni.lr.fit)
round(exp(pni.lr.fit$coefficients), digits = 2)
round(exp(confint(pni.lr.fit)), digits = 2)

```

###Interpretation of the results:
Occult metastasis ~ Perinural Invasion
<br/>
OR = 2.35 (95% CI 1.08 - 5.26)
The odds of developing occult metastasis when perinural invasion is present is 2.35 times the odds for those who do not have perinural invasion
<br/>
This result is  statistically significant as the null value of 1 is not included in the 95% CI.

## Univariate logistic regression  - Model 4
<br/>
1) Predicting probability of occult metastasis in lymph nodes
<br/>
2) risk factor - Level of differentiation
```{r Degree of differentiation model, echo=FALSE}
#Grade of differentiation
#table(baseline_scc_doi$biopsy_differen, exclude = NULL)
DD_log.reg<-baseline_scc_doi%>%
  select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, peri_neu_inva,biopsy_differen,n_patho_stage, n_clinc_stage)%>%
  filter(n_clinc_stage== "N0" & n_patho_stage != "Not Applicable", !is.na(biopsy_differen))%>% view()
#110 participants for this model

DD_log.reg$occultmet<-NA
for (k in 1:nrow(DD_log.reg)) {
  if (DD_log.reg$n_patho_stage[k] != DD_log.reg$n_clinc_stage[k]){
    DD_log.reg$occultmet[k]<- 1 #meaning yes occult metastasis seen
  } else{
    DD_log.reg$occultmet[k]<-0 #meaning no occult metastasis
  }
}  

#table(DD_log.reg$occultmet, DD_log.reg$biopsy_differen, exclude = NULL)
#class(DD_log.reg$biopsy_differen)
DD_log.reg$biopsy_differen<-factor(DD_log.reg$biopsy_differen, levels=c("Poorly","Moderate","Well"))
table(DD_log.reg$occultmet, DD_log.reg$biopsy_differen, exclude = NULL)

DD.lr.fit<-glm(occultmet~biopsy_differen, data = DD_log.reg, family = binomial)
summary(DD.lr.fit)
round(exp(DD.lr.fit$coefficients), digits = 2)
round(exp(confint(DD.lr.fit)), digits = 2)

```

###Interpretation of the results:
Occult metastasis ~ Degree of differentiation
<br/>
1) Moderately differentiated
OR = 0.69 (95% CI 0.19 - 2.66)
The odds of developing occult metastasis when Moderate degree of differentiation is present is 0.69 times the odds for those poorly differentiated tumors.
<br/>
This result is not statistically significant as the null value of 1 is included in the 95% CI.
<br/>
<br/>
1) Well differentiated
OR = 0.26 (95% CI 0.06 - 1.13)
The odds of developing occult metastasis when well differentiated is 0.26 times the odds for those poorly differentiated tumors.
<br/>
This result is not statistically significant as the null value of 1 is included in the 95% CI.


## Univariate logistic regression  - Model 5
<br/>
1) Predicting probability of occult metastasis in lymph nodes
<br/>
2) risk factor - Lymphovascular invasion
```{r , echo=FALSE}
# baseline_scc_doi%>%
#   select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, n_patho_stage, n_clinc_stage)%>%
#   filter(n_clinc_stage== "N0")%>%
#   count(n_patho_stage) #leaves us with 82 participants


LVI_log.reg<-baseline_scc_doi%>%
  select(Patient.IDID.du.patient, surgery_type,doi_cat,doi_125_round, vas_lymph_inv,n_patho_stage, n_clinc_stage)%>%
  filter(n_clinc_stage== "N0" & n_patho_stage != "Not Applicable", !is.na(vas_lymph_inv))%>% view()
#127 participants for this model

LVI_log.reg$occultmet<-NA
for (j in 1:nrow(LVI_log.reg)) {
  if (LVI_log.reg$n_patho_stage[j] != LVI_log.reg$n_clinc_stage[j]){
    LVI_log.reg$occultmet[j]<- 1 #meaning yes occult metastasis seen
  } else{
    LVI_log.reg$occultmet[j]<-0 #meaning no occult metastasis
  }
}  

table(LVI_log.reg$occultmet, LVI_log.reg$vas_lymph_inv, exclude = NULL)

LVI.lr.fit<-glm(occultmet~vas_lymph_inv, data = LVI_log.reg, family = binomial)
summary(LVI.lr.fit)
round(exp(LVI.lr.fit$coefficients), digits = 2)
round(exp(confint(LVI.lr.fit)), digits = 2)

```
###Interpretation of the results:
Occult metastasis ~ Lympho-vascular invasion
<br/>
OR = 2.50 (95% CI 1.07 - 5.85)
The odds of developing occult metastasis when lympho-vascular invasion is present is 2.50 times the odds for those who do not have lympho-vascular invasion
<br/>
This result is  statistically significant as the null value of 1 is not included in the 95% CI.


## Univariate logistic regression  - Model 6
<br/>
1) Predicting probability of occult metastasis in lymph nodes
<br/>
2) risk factor - DOI - categorical variable

```{r, echo=FALSE}
class(log.reg$doi_125)
log.reg$doi01_cat<-cut(log.reg$doi_125, 
    breaks = c(0,1.9,3.9,5.9,7.9,9.9,50) ,
    labels = c("0-1.9","2-3.9","4-5.9","6-7.9","8-9.9",">=10"),
    include.lowest = T)

table(log.reg$doi_125, log.reg$doi01_cat, exclude = NULL)

doicat1.lr.fit<-glm(occultmet~doi01_cat, data = log.reg, family = binomial)
summary(doicat1.lr.fit)
round(exp(doicat1.lr.fit$coefficients), digits = 2)
round(exp(confint(doicat1.lr.fit)), digits = 2)

```


```{r, echo=FALSE}

log.reg$doi02_cat<-cut(log.reg$doi_125, 
    breaks = c(0,2.9,4.9,6.9,8.9,10.9,50) ,
    labels = c("0-2.9","3-4.9","5-6.9","7-8.9","9-10.9",">=11"),
    include.lowest = T)

table(log.reg$doi_125, log.reg$doi02_cat, exclude = NULL)

doicat2.lr.fit<-glm(occultmet~doi02_cat, data = log.reg, family = binomial)
summary(doicat2.lr.fit)
round(exp(doicat2.lr.fit$coefficients), digits = 2)
round(exp(confint(doicat2.lr.fit)), digits = 2)


```

